 \FloatBarrier
 
\citet{goodman1967age} offers a suite of
formulas to determine the stable age-sex composition of a population taking into
account the vital rates of both sexes, assuming one can assign a relative weight
(summing to 1) to male and female fertility. This model was presented in
Section~\ref{sec:googmanage} for age-structured populations, and will now be
translated for the case of remaining-years structured populations. Recall that
this model entails two trade-offs: 1) one must (arbitrarily) choose dominance
weights, and 2) these weights are constant. The fact of having constant weights 
keeps the solution linear (interaction-free), but less realistic. The final result is bracketed by the
cases of male and female dominance, but the gap between these two extremes 
also measures the demographer's subjective leeway, which we would like to
minimize. Both of these drawbacks may be reduced in the case of $e_y$-structured
populations, since: 
\begin{enumerate}
  \item $e_y$-structured populations have a more stable (in terms of
  distributional variation from year-to-year) structure than age-structured
  populations.
  \item Mate-selection with respect to remaining years of
life is nearly random in $e_y$-structured populations (see
Section~\ref{sec:exobsexpected}).
  \item The difference between male and female dominance (in terms of projected
birth counts) is often reduced, thereby limiting of the impact of the
demographer's ``dominance caprice'' on results (See
Section~\ref{sec:exdivergence}).
\end{enumerate}
Points (1) and (2) reduce (but do
not eliminate) the necessity of sex-interactions in a model. By this it is meant
that the proportional difference in results from one choice of model weights
over another is simply diminished. This being so, the comparative advantage of a
more sophisticated or realistic model is to some degree diminished. Since the
weighting coefficients in this model do not change, we have taken the extra step
to design a projection matrix for this dominance model, and we put this to use
to study some of the transient properties of the present model, as well as
examine the resulting stable population structure.
