 \FloatBarrier
The formal relations presented in
Sections~\ref{sec:2sexlinearmain}~and~\ref{sec:2sexlinearother} establish
coherence, and some merits have been presented. This section offers tools
more relevant to the discrete world of applied demography. The model contained
in Equation~\ref{eq:lineartwosexrenewal} is best replicated with a projection
matrix, similar in concept to that offered for the single-sex $e_x$-structured
case in Section~\ref{sec:ex1sexleslie}. The two-sex linear projection matrix combines
the projection of each sex jointly in a single instrument, the construction of 
which is more involved than the single-sex case: Four times more involved to be precise. 

Assuming $n$ $e_x$-classes each for males and females, the dimensions of the
present matrix will be $2n \times 2n$, where male and female
$e_x$-classified population vectors by be joined, for instance end-to-end in a
single population vector. The convention used in this description will place
males, ordered by remaining years of life in positions $1:n$ of the vector
$\textbf{p}$ and females ordered by remaing years in positions $(n+1):(2n)$ of
$\textbf{p}$, i.e. end-to-end. This being so, the projection matrix $\textbf{Y}$
must conform with these locations of males and females, locating survival and 
fertility appropriately.

$\textbf{Y}$ is divided into four main blocks. The top left block is
nearly identical to the male single-sex case, and the bottom left block is
nearly identical to the female single-sex case. Both of these two submatrices
contain survival (all 1s) in the superdiagonal. Fertility is analagous, but must
be weighted according to $\sigma$ ($1-\sigma$). The lower left submatrix
contains $M-F$ fertility weighted by $\sigma$ and distributed over female death probabilities,
$d_a^f$, and the upper right matrix contains $F-M$ fertility, weighted by
$1-\sigma$ and distributed according to male death probabilities. As with the
single-sex projection matrix, fertility destined to die in the first year of
life must be further discounted by $\lambda$ to account for the part of infant
mortality that occurs before Dec. \nth{31} of the calendar birth year. In this
case $\lambda$ may optionally be entered separately for males and females.

% need to adjust spacing big-time!
\begin{landscape}
\begin{matrix}[h!]
\caption{A full two-sex remaining years ($e_y$)-structured
projectionmatrix, $\textbf{Y}$} 
\label{matrix:ex2sex}
 \begin{adjustwidth}{-.2in}{-.5in}
 \centering
\footnotesize{
$\bordermatrix{
  {e_y }    \vspace{1.2em} & 0_t^M                                             &
  1_t^M                                       & 2_t^M                                    & 3_t^M   & 0_t^F                                                & 1_t^F                                          & 2_t^F                                       & 3_t^F  \cr 0_{t+1}^M \vspace{1.2em} & \sigma  (1-\lambda^M)\tfrac{f_0^{M-M}d_0^M}{2}       
  & \sigma (1-\lambda^M)f_1^{M-M}d_0^M + 1          & \sigma (1-\lambda^M)f_2^{M-M}d_0^M           &  0      & (1-\sigma)(1-\lambda^M)\tfrac{f_0^{F-M}d_0^M}{2}         & (1-\sigma)(1-\lambda^M)f_1^{F-M}d_0^M             & (1-\sigma(1-\lambda^M)f_2^{F-M}d_0^M &  0     \cr 
  1_{t+1}^M \vspace{1.2em} & \sigma \tfrac{f_0^{M-M}d_1^M}{2}                  &
  \sigma f_1^{M-M}d_1^M                       & \sigma f_2^{M-M}d_1^M + 1                &  0      & (1-\sigma) \tfrac{f_0^{F-M}d_1^M}{2}                 & (1-\sigma)f_1^{F-M}d_1^M                      & (1-\sigma)f_2^{F-M}d_1^M                    &  0     \cr 2_{t+1}^M \vspace{1.2em} & \sigma \tfrac{f_0^{M-M}d_2^M}{2}                  &
  \sigma f_1^{M-M}d_2^M                       & \sigma f_2^{M-M}d_2^M                    &  1      & (1-\sigma) \tfrac{f_0^{F-M}d_2^M}{2}                 & (1-\sigma)f_1^{F-M}d_2^M                      & (1-\sigma)f_2^{F-M}d_2^M                    &  0     \cr 3_{t+1}^M \vspace{1.2em} & \sigma \tfrac{f_0^{M-M}d_3^M}{2}                  &
  \sigma f_1^{M-M}d_3^M                       & \sigma f_2^{M-M}d_3^M                    &  0      & (1-\sigma) \tfrac{f_0^{F-M}d_3^M}{2}                 & (1-\sigma)f_1^{F-M}d_3^M                      & (1-\sigma)f_2^{F-M}d_3^M                    &  0     \cr 0_{t+1}^F \vspace{1.2em} & \sigma (1-\lambda^F)\tfrac{f_0^{M-F}d_0^F}{2}     &
  \sigma (1-\lambda^F)f_1^{M-F}d_0^F          & \sigma (1-\lambda^F)f_2^{M-F}d_0^F       &  0      & (1-\sigma) (1-\lambda^F)\tfrac{f_0^{F-F}d_0^F}{2}    & (1-\sigma)(1-\lambda^F)f_1^{F-F}d_0^F +1          & (1-\sigma)(1-\lambda^F)f_2^{F-F}d_0^F           &  0     \cr 1_{t+1}^F \vspace{1.2em} & \sigma \tfrac{f_0^{M-F}d_1^F}{2}                  &
  \sigma f_1^{M-F}d_1^F                       & \sigma f_2^{M-F}d_1^F                    &  0      & (1-\sigma) \tfrac{f_0^{F-F}d_1^F}{2}                 & (1-\sigma)f_1^{F-F}d_1^F                      & (1-\sigma)f_2^{F-F}d_1^F +1                 &  0     \cr 2_{t+1}^F \vspace{1.2em} & \sigma \tfrac{f_0^{M-F}d_2^F}{2}                  &
  \sigma f_1^{M-F}d_2^F                       & \sigma f_2^{M-F}d_2^F                    &  0      & (1-\sigma) \tfrac{f_0^{F-F}d_2^F}{2}                 & (1-\sigma)f_1^{F-F}d_2^F                      & (1-\sigma)f_2^{F-F}d_2^F                    &  1     \cr 3_{t+1}^F \vspace{1.2em} & \sigma \tfrac{f_0^{M-F}d_3^F}{2}                  &
  \sigma f_1^{M-F}d_3^F                       & \sigma f_2^{M-F}d_3^F                    &  0      & (1-\sigma) \tfrac{f_0^{F-F}d_3^F}{2}                 & (1-\sigma)f_1^{F-F}d_3^F                      & (1-\sigma)f_2^{F-F}d_3^F                    &  0 }$}
  \end{adjustwidth}
\end{matrix}
\end{landscape}
Matrix~\ref{matrix:ex2sex} is a schematic representation of a two-sex
$e_x$-structured projection matrix. This example contains four classes of life
expectancy in order to economize space (as opposed to the six shown in
Matrix~\ref{matrix:ex1sex}). Such a matrix amenable to the data used in
this thesis would have final dimensions $222 \times 222$, since we work here
with 111 remaining life expectancy classes\footnote{The 111 $e_x$ classes
are namely derived from the 111 ages of $d_x$ provided by the HMD, ages
$0-110+$.}.

