\FloatBarrier

The purpose of this chapter was to provide intuition into the nature of the
two-sex problem by means of data-driven illustation. Males and females differ in
the age patterns and levels of all demographic phenomena. These differences
may partially reflect differences in the evolutionary optimizaton of the life
course, such as the sex ratio at birth and male excess mortality offsetting each
other. The magnitude of the effect of these two factors varies over time, but
the sign has proven consistent, at least for the two populations explored.
Fertility effects have been shown to be far less consistent. While differences
in the reproductive span, the age-bounds to reproduction, may also be attributed
to the relatively slow evolution of the life course, different locations on the
repective male and female marginal distributons are evidently maleable in
contemporary societies. Thus, we may observe inconsistent influence from the
shape of fertility on male versus female growth rates-- This we note was
especially the case for the Spanish population. Further, differences in the
overall level of fertility, which are separable from shape effects, will 
owe primarily to differences in male and female exposure levels, and hence will vary
from year to year depending on population structure, which is itself the fruit
of all manner of past demographic phenomena.

It is for this reason that fertility (marriage) balancing has been the primary
focus of methods intended to account for the two-sex problem in demography.
Fertility is the source of new generations in population models, i.e. the
starting point in a population model. One may conceivably, and will virtually
always do in ractice, conceive of male and female mortality as mutually
exclusive forces. Therefore, once a new cohort is produced in a population
model, the rest is taken care of by the respective sex-specific mortality
schedules. This leaves the sex ratio at birth and fertility to be
thoughtfully dealt with in models, and this will be the topic of the following
chapter.
