% 
This dissertation would not have been possible without the continued support of
the Centre d'Estudis Demogr\`{a}fics, which was my academic home for the five
years from 2008 until 2012. Thanks are owed to many of the faculty and
staff of this wonderful institution. Most especially I thank my director since
the start, Dr. Albert Esteve, who gave me all the space, encouragement and
support I needed during my doctoral studies -- you kept me pointed in the right
direction. Thank you to Anna Cabr\'{e} for welcoming me to the CED,
for first imparting me with the fundamentals of classical demography, and for
having been a constant source of wisdom and advice. Thanks to I\~{n}aki
Permanyer for helping me with some methodological issues and an overall critique. 
Thank you Daniel Devolder for always
being available to assist, for letting me pilfer your library, and for always
understanding what it was I was trying to say. Thanks to my CED peers for your
commraderie and empathy. Thank you Soco for keeping me legal and in
order.

Thank you to Vladimir Canudas Romo, for your encouragement and
empowerment. I would never have undertaken a formal demographic topic were it
not for you, and I would have never landed in Berkeley were it not for
you. Thanks to Alexia F\"{u}rnkranz-Prskawetz for your addictive enthusiasm for
models. Thank you Adrien Remund and Felix R\"{o}ssger for keeping me on my
toes -- more and more I think, thanks to you guys, that there is plenty of
low-hanging fruit left out there to pick.

Thank you John Wilmoth for bringing me to Berkeley, for your generosity, trust
and continuous positive feedback. Thanks to Magali Barbieri for your
encouragement and empathy, to Carl Boe for being available 24 hours to help with
methodological issues of any scale and relevance to the task at hand. Thank you
Robert Chung for sharing your experience, for your dedication of time
and mental resources to the problems that I have made, and for your goodwill in
general. Thank you to the HMD as a whole for providing me with an at once demanding,
didactic, trusting, and friendly team and environment.

There are also acts of random kindness that possibly happen in all academic
communities, but which we may as well recognize here with names: Griffith Feeney
scanned and sent me his PhD thesis amidst an intercontinental move -- wow,
thanks! Ron Lesthaeghe also dug up, photocopied, and sent me via snail mail a copy
of Stanlye Wijewickrema's thesis at no cost -- wow, thanks! Robert Chung, you
also deserve mention in the good samaritan section for helping me
``just-because.''

Thank you to my mom and dad for your faith in me and for your always genuine
curiosity about what it is I do. You've invested heavily in my education, and never once
faltered in light of my academic caprice. This has been a source of emotional
stability in light of my otherwise seemingly haphazard bet-taking, which finally
appears to be paying off.

Thank you Ainhoa for always believing in me, for telling me so, and for keeping
me pointed straight. This dissertation would never have been finished without
you!

My stay in the CED was financed primarily by a
fellowship from the Ag\`{e}ncia de Gesti\'{o} d'Ajuts Universitaris i de Recerca
of Catalunya, and for that I owe a debt of gratitude to my metahost, Catalonia,
and to Soco and Albert for help with the entailed paperwork, which I
would have never otherwise been able to manage. Other sources of funding
included an in-house stipend from the CED in 2008; a mobility scholarship 
from the Ministerio de Educaci\'{o}n of Spain, which in principle supported 
part of my stay in Lund for the EDSD; a supplemental stipend from the
MPIDR during the start of the EDSD in Rostock in 2009; and especially the
WORLDFAM grant from Albert Esteve, which funded all of my conference
participation during the course of my PhD studies, as well as summer courses at
the ICPSR in 2009. The final phase of my financial support, after my PhD fellowship, came from my current employer, the Department of
Demography at the University of California, Berkeley, where I currently work for
the Human Mortality Database project, but where I also received constant encouragement to finish my PhD and
sizeable chunks of on-the-clock time alotted to work exclusively on this
dissertation. More than any government or institution, my mom
and dad are responsible for the material investment in my education that got me this far,
including occassional injections of support throughout the doctoral process.






