Demographers study population structure and demographic flows. In order
to assess the magnitude and intensity of demographic phenomena in real-world
populations, one must first remove potential distortions introduced by
population structure-- but population structure is itself an outcome
of demographic phenomena. Here, demographic phenomena refer to natality and
mortality, and population structure refers to classifying information such as
time and sex. There are other structuring variables whose effects we would also
like to purge were the measurement of demographic phenomena 
our primary objective, and there are other kinds of phenomena that would also
need to be measured if the analysis of structure were the primary
objective. Such description is of secondary importance in this dissertation. Our
objective is to study an abstraction of population processes, namely the 
renewal model for closed populations structured by sex and time.

That we are concerned with the role of both sexes in the modeling of population
renewal should be no surprise, as humans reproduce sexually. That it is a
challenge for models to incorporate information from both sexes in the modeling
of fertility (marriage, reproduction) has been firmly established since
\citet{karmel1947relations}. This task is challenging because models must
produce a single result, a fertility outcome, from two information sources
(males and females), which when handled apart produce incongruous results. 
There is no obviously correct way to achieve this balancing act, although a
large number of suggestions have been made. We typically call these suggestions ``solutions'',
but they are not solutions in the sense of a solution to a math problem. A
solution in the context of the present problem simply means that a reasonable
result is produced in accordance with a predefined set of modeling objectives
decided upon either by the demographer or consensus. The problem has not been
(and may never be) solved in the sense of a necessary and best model. Instead,
solutions are weighed in terms of fulfilling desirable properties versus theoretical 
or practical parsimony.

In this dissertation, we deal only with two population subgroups: males and
females, each of which is structured by some notion of time. The modeling
problems that emerge when dealing simultaneously with the two sexes may be
conceived of as a minimal case of the much larger practical problems 
presented by modeling populations subdivided into $N$
groups. Applied demographers often simultaneously project
populations divided into regions, races, educational groups, and a potentially
large number of other categorical distinctions. The modeling
challenges presented by multigroup plurality may in this sense be thought of as
the general problem, within which the two-sex problem is just a particular
instance that must be dealt with under particular constraints. While the two
sex problem be in a sense be reduced to the notion of the $N$-group problem,
the sexes interact in a way, and reproduction is constrained in a way, that does
not pertain to other kinds of population subgroups-- There is namely no
``race-ratio'' or ``state-ratio'' at birth akin to the sex ratio at birth.

Later in the
present work we will provide some measure of the magnitude of the two-sex problem, 
and it will be concluded that the
magnitude is large enough to be worth thoughtfully accounting for in
measurements and modeling. Were models to encompass even more groups, the potential
discrepancies entailed by simultaneous modeling would be even larger. That we may arrive at
insights from the more familiar two-sex case that apply to the $N$-group case is
a further motivation for thoughtful exploration of the problem at hand.
$N$-group generalizations will not be explored in the present work, though with
some additional work, solutions discussed here may be extended in this
direction.

The balancing of the sexes in models without considering age is
much simpler both conceptually and in practice, as it is just a matter of
choosing some middle ground between males and females. Most of the literature on
the two-sex problem, and the properties that demographers deem desirable in
solutions, deals with the time-structuring variable, age. That modeling
decisions must be made both with respect to the interaction between sexes and
the interaction between ages makes the problem an altogether complex one. 

What is age but time passed since birth? Thus, age is \textit{time} with respect
to one of the demographic phenomena that we incorporate in models of population
renewal. The reason why demographers care about age is because all
demographic phenomena vary by age in known ways, and so in order to measure the
pure force of a demographic phenomenon one does well to take age into account-- the age patterns of
demographic phenomena exhibit empirical regularities that lend themselves to
modeling \citep{coale1996development}. Age counts 
up from birth, starting at the beginning. We measure milestones and the lifecourse in 
terms of age; statistics are collected by age or year of birth, and age is in 
short \textit{known}. We do not know when we will die, but
this is also something that demographers think on. Namely, in properly
accounting for age (time since birth), we may faithfully approximate death
probabilities for each age, and therein know something about our probable time
of death. This later question is a subject of considerable interest!

Demographers, and especially actuaries, regularly think about, estimate, a
probable time of death for persons of a particular ages-- that were born in
particular years. Might we not also venture to take things a step further?--
What if age were counted down to death instead of up
from birth? Literally, what might we learn about demographic phenomena and
population structure if beyond age (and due to the information we glean from
age) we were to structure populations by sex and remaining years of life? That
is a big job, and we will fail to complete it, instead laying out only the groundwork
for population renewal models wherein age is exchanged for remaining years of
life.

It is my stance that population renewal models ought to account for both sexes,
and for this reason roughly equal attention is given in this dissertation to the
two-sex problem-- a problem that never goes away-- and population structured
by remaining years of life-- a somewhat novel concept that must be hashed out
before again complicating things with the two-sex problem. I apologize for any lack of
rigour on both fronts. Namely, I neither reproduce formal proofs for the
properties of the solutions that I treat, nor do I provide proofs for the
(many) claims that I make. The filling of this gap is left for a later day--
either someone will do the favor of proving my claims right or wrong or I
will find the time to learn to do so. Instead we are lead in this dissertation
primarily by intuition, and I have placed a premium on the
data-grounded demonstration of the methods I propose. After all, might we not
wish to free these formal demographic musings from the vacuum and see what might
be learned? There is therefore the risk that some conceptual error or
miscalculation of mine-- and all errors and miscalculations herein are mine
alone-- will be a setback. This is my risk alone, but the possibility is
not that distressing. Rather it is inherent to the business of charting new
territory, and this I have every pretension of doing. If the maps I draw are no
good, the territory explored may still be good.

So it is that sex and \textit{time} are the structuring variables of
interest in this dissertation. Aside from comparing two-sex models (and often the male
and female one-sex models), we at times compare models that
specify age with models that specify time until death. We will learn that the
specification of time in models has a large impact on results, and it could
be said (with a pinch of jest, of course) that we herein expose a
fundamental \textit{two-age} problem in demography. This was of course not our goal from the
outset of the doctoral process. The narrative of how this dissertation came to
take the shape that it has is as follows:

The original objective for this dissertation was to hash out a survey of two-sex
``solutions'' and implement them in a standard and reproducible format while
applying each to contemporary datasets. I chose the topic after completing the
EDSD in the summer of 2010 in order to force myself to improve my formal
demographic skills, which I had only begun to develop in that program. That
is, I knew it would be difficult and possibly beyond my abilities. And so I
began at the beginning, collecting all the materials I could locate on the 
two-sex problem, and reproducing methods in no particular order. After a few 
successful attempts (reproducing \citet{schoen1981harmonic},
\citet{mc1975models} and \citet{henry1972nuptiality}) on ad hoc acquired data, I
came to realize that all the methods in my scope will essentially require or
the same input data-- basic exposures, and births cross-tabulated by
sex and age of father and mother-- so I diverted attention to standardizing some
datasets to use throughout this dissertation-- The US and Spanish populations
for about the past four decades. I decided to discard or translate methods
dealing with the two-sex problem in marital transitions, in the first place 
because Spain lacks good estimates of marital status exposures, 
and these would need to have been produced artisanally (laboriously). This
choice reduced the implementation workload, but the sex balancing strategies
from analytical family demography have still been taken into consideration where
possible.

When the mathematics or
presentation style in a given article were over my head, I typically took a few
steps back to some earlier or less complex method, or alltogether went back to the basics 
in \citet{sharpe1911problem}, \citet{kuczynski1932fertility}, \citet{coale1972growth} 
or \citet{caswell2001matrix}. Some methods that were beyond my grasp in the 
begining \citep[e.g.][]{mitra1978derivation, gupta1978alternative} were finally
understood and implemented later down the road. Others I still do not 
understand\cite[e.g.][]{choo2006estimating}, despite having
reproducible code!

All along I had no vision or pretense of designing a new method, but I rather
na\"{\i}vely assumed that gradual familiarity with the tools at hand would lead
me to some minor tweak or meaningful critique of the existing palette of methods at
hand. For two years I did not produce anything novel, and only managed to
reproduce a few branches of the above-mentioned survey of methods, and my
resolve waned. A spontaneous conversation with John MacInnes sparked
what was to become the second, but dominant axis of this thesis, our realignment 
of age in renewal models. John
mused, namely, what a population pyramid would look like if it were drawn with
life expectancy on the vertical axis instead of age. We tried to imagine what
shape such a pyramid would assume, but were on the whole left guessing. I took a
stab at how one might go about transforming age-classified population counts to
remaining years-classified population counts, and came up with what is here
Equation~\ref{eq:dxredist}. Later I realized that the central component to that
formula, which says ``what is the probability of dying at age $x+n$ given
survival until age $x$'', is fairly fundamental and already lying around in
formal demography, probably in various texts and minds-- I spotted it in the
wild in \citet{vaupel2009life}, and more prominently in
\citet{miller2001increasing}, but it's certainly well-known and/or easily
derivable. I have never seen this formula applied to redistribute population counts in the manner
suggested here, although the concept of remaining years until death has
certainly been considered. Miller refers to this
temporal concept as \textit{thanatological} age\footnote{This phrase does not
appear in the cited paper, but Miller informs me that the phrase was
coined by Ken Wachter.}.

Shortly thereafter, after looking at many such remaining years pyramids and
coming to some exciting conclusions, I realized that one may just as well
restructure \textit{any} age-classified data in the same way. So I took a 
look at some thusly restructured fertility rates, explored
some more, and spontaneously resolved to try to figure out what form the
fundamental Lotka equations would obtain if reworked to be based on
remaining-years classified data. From that moment I was self-obliged to bring \textit{this} family of
population models to bear upon the two-sex problem. Is the problem the same?
Will this transformation teach us anything? Does it make sense to simply
project remaining-years structured populations alongside age-structured
populations? That is what this dissertation is about.

As hinted before, it is the case that when one estimates population growth (or
some other interesting quantity) from a remaining-years-classified population,
one arrives at a different result than that derived from the analogous
age-structured model. I do not undertake the worthy task of formalizing the
difference between age and remaining-years structured growth rates, as did
\citet{karmel1947relations} for the difference between male and female
reproduction rates (or \citet{coale1972growth}, put differently). As for this
new discrepancy, I have only managed the less glamorous feat of pointing out
that it exists, as did \citet{kuczynski1932fertility} for the male
and female cases. There is plenty of work left to do, and it is my hope herein
to stimulate discussion in the discipline of demography about whether structuring
aggregate demographic data (and the models derived from these) by remaining
years of life may contribute further insights into human population dynamics. At
times in this dissertation, I will posit how the nature of
remaining-years-structured populations is agreeable to stable population theory
and lends itself to population projections in general and to the sex balancing
undertaken in two-sex solutions in particular. Further, the strategies that
demographers have developed to patch the two-sex problem will provide us with
insights into the new discrepancy presented by our restructuring of age.
