Before using the term at length, we offer a quick definition of the two-sexproblem:\begin{singlespace}\begin{quote}The \textbf{two-sex problem}: Separate male and female predictions of eventswill differ when 1) the events depend on both sexes, 2) the same events form the numerators of the rates of each sex that are used in the prediction,and 3) the exposures used to calculate these same rates change from year to year(time to time). We would like a single prediction based on information from thedemographic rates of each sex.\end{quote}\end{singlespace} Such events include births and heterosexual marriages. A manifestation of this problem in formal demography is that the male and female single-sex renewal models calculated from the data of any observed  population will diverge from one another, and therefore neither represents the population as a whole. Models that account for the rates of both sexes so as to produce a single and consistent result or prediction are variously referred to throughout this dissertation as two-sex \textit{methods}, two-sex \textit{solutions}, two-sex \textit{rate balancing}, two-sex  \textit{adjustments}, and so on with no strict distinction between terms.  The two-sex problem in human demography has until this work been defined  and studied either for populations that are structured by sex only, for populations structured by both sex and age, or for populations structured by sex, age, and marital status. The problem enjoys a long history in formal demography, and mostcontemporary applied demographers likely have a rough idea of what the problemis, but it is not at the forefront of the consciousness of most contemporarypracticing demographers. For this reason we dedicate a sizable portion of this dissertation, Part 1, to defining, describing,and decomposing (Chapter~\ref{ch:Measuring}) the two-sex problem for populationsstructured by age and sex, and to describing a set of methods developed to dealwith this problem (Chapter~\ref{sec:modelingapproaches}). The primary (and likelyoriginal) contribution of Chapter~\ref{ch:Measuring} is the decomposition of the sex-gap in intrinsic growth rates into the various vital rate components of the classic renewal model. This effort will convince the reader that the two-sexproblem is worth thinking about and accounting for, and it justifies a largeportion of the remainder of this dissertation.These activities are carried out formally in the context of the fundamentalLotka equations for single-sex population renewal and empirically on the basis US and Spanish data from the past four decades. The choice to work with the two-sex problem inrenewal models is not an idle endeavor, as the practice of projectingpopulation -- the bread and butter of applied demographers -- is grounded inmodels of population renewal. That is to say, in implementing a population projection, one always has a renewal model in mind, and all indices ofpopulation reproductivity refer to a notion of renewal. The methodsdescribed herein are applicable as-is to implementation in populationprojections.We suppose that part of the reason that practicingdemographers often do not account for the two-sex problem in projections (and elsewhere) is that formal demographers have been unable to supply aconsensus solution on how to deal with it. Of course, the lack of consensus ondemographic methods is no obstacle for demographers in other tasks -- there are different ways to calculate lifetables, estimate exposures, smooth or graduatedemographic schedules, and so forth, but this does not stop demographers fromdoing these things. Unfortunately there is nobest and true method to account for the two-sex problem in renewal models or projections; there are only approximations, and this is likely to remain the case. For thisreason, in Chapter~\ref{sec:modelingapproaches} we present a suite of approaches applicable to the US and Spanish example data used throughout this work. These methods are evaluated in situ. The selection oftwo-sex methods examined is not exhaustive, but has been selected onan ad hoc basis of convenience to the author. It is hoped that the set oftwo-sex model implementations provided here will be translatable andtransferible to population projection engines. In this way, the demographer maythoughtfully select from among the two-sex methods conscientiously, just asevery maker of lifetables chooses a method to calculate $a_x$.For each method separately, we begin by explaining the model of renewal. Thisis followed by a step-by-step guide to estimating the intrinsic growth rate,$r$, for the given model. All such estimation procedures have been developed bymodifying the fast-converging method of \citet{coale1957new} to the renewalmodel at hand. At times other parameters of the stable population are also presentedon the basis of $r$. In all cases, some empirical results are derived for themethod in question and some model properties are discussed. Some of thesemethods will be modified from their original context to suit the needs of thisdissertation.The first two methods presented, those of \citet{pollard1948measurement} inSection~\ref{sec:pollardage} and\citet{mitra1978derivation} in Section~\ref{sec:mitraage}, serve more for context and curiosity than forpractical use. These two models namely do not distribute the fertility of eachsex of birth between both sexes of parents, which either introduces instabilityor an unrealistic notion of renewal. Section~\ref{sec:googmanage} presents oneway of using \textit{dominance weights} to decide how to divide the information source for fertility prediction between males and females. We find this methodconvenient, and so it is translated to the case of populations structured bythanatological age later on. In the following we explore three models that make use of the full jointdistribution of fertility by the ages of both parents. The model of\citet{gupta1978alternative} is presented in Section~\ref{sec:dasgupta}. This isfollowed in Section~\ref{sec:ageharmonic} by a method to balance fertilitypredictions by taking a mean of male and female exposures in each age combination, which we largely demonstrate on thebasis of the harmonic mean, although the mean could be of any kind. This methodis attractive for different reasons, and so will also be translated to the caseof remaining-years structure. Finally, Section~\ref{sec:IPF} presents aniterative method to balance the birth counts predicted by males and females indifferent age combinations, and this method is also implemented for the case ofremaining-years structured populations. Section~\ref{sec:other} very brieflydiscusses a large number of other modeling approaches that have also beentaken, or that could be taken, with respect to two-sex population modeling, butthat are beyond the scope of this dissertation.The two-sex problem has never been explored for the case of populationsstructured by remaining years of life (thanatological age). This is necessarily so because Lotka's single-sex renewal model \citep{sharpe1911problem} has neverbeen translated to the case of populations structured by thanatological age, and this latter endeavor banks on structuring a population by thanatological age in the first place, which is apparently itself novel. Thislast item, structuring populations by remaining years of life, is done by meansof a slight modification to formulas that already exist, and so most of thenovelty therein owes to framing the pre-existent desire for the remaining-years perspective interms of population structure. These will be the first tasks of Part 2of this dissertation.Section~\ref{sec:thetransformation} provides and relates the formulas totranslate age-structured data into remaining-years structured data. The mostbasic demographic data amenable to such restructuring are population counts.This we present in Section~\ref{sec:expopstruct1} for our two populations.Many of the results that fall out of this activity (or consequences of it) areof potential immediate utility, and so we briefly discuss some aspects ofthanatologically structured populations, such as uncertainty(Section~\ref{sec:structuncertainty}), the incorporation of mortalityimprovements in the method (Section~\ref{sec:eximprov}), and a couple ofthe more obvious and direct measures of population \textit{aging}(Section~\ref{sec:exageing}).A further conceptual leap is necessary before the business of populationmodeling may be undertaken atop this new population structure -- namely, theobservation that this method of population restructuring works just the same for any age-classified data. Ofrelevance for our population modeling objective is the restructuring ofage-classified birth counts and thencefrom the production of fertility rates byremaining years of life. Section~\ref{sec:exrates} explores these fertilityrates for the US and Spanish populations. The practice of population modelingon the basis of such rates will be palatable only if a high degree of empiricalregularity is present in the data. This we confirm, revealing for the firsttime the characteristic shape of thanatologically-structured fertility rates--the thanatological analogue to ASFR, which we call $e$SFR. Such rates may bepresented in bulk in the form of a standard ``\textit{remaining years} $\times$year'' demographic surface, but we also summarize the intensity of fertilityfor a given year in Section~\ref{sec:etfr} by summing rates over remaining years of life to producethe thanatological analogue to TFR, $e$TFR, which enjoys thesame basic interpretation as TFR yet sums to a different value. InSection~\ref{sec:exdivergence} we demonstrate that male and female predictionsof future births based on thanatological rates will tend to differ by less than is the case for predictions made on the basis of age-structured rates. Lastly,also as a brief diversion, we relate $e$TFR to a remaining-years reformulationof Fisher's reproductive value.Having demonstrated sufficient empirical regularity in the remaining-yearspattern to the necessary vital rates, we reconceive of the classic notion ofage-structured population renewal in Section~\ref{sec:exrenewal}. Thatdescription is intended to be intuitive, and we hash out a diagram(Figure~\ref{fig:exrenewal}) that should serve as a visual mnemonic for the concept of renewal in populations structured by remaining years of life. InSection~\ref{sec:ex2sexequation} we undertake a broad project of defining thesingle-sex renewal equation(s) for this perspective of population, including theprovision of an iterative method to optimize $r$(Section~\ref{sec:exrenewalit}), formulas for the derivation of stablepopulation structure (Section~\ref{sec:ex1sexother}), and induction of theprojection matrix\footnote{This product is the analogue to the well-known Lesliematrix.} that corresponds to the remaining-years perspective (Section~\ref{sec:ex1sexleslie}). After displaying some trends inremaining-years $r$ for the US and Spain (Section~\ref{sec:trendsinrex1sex}), we explore the speed of divergence between the single-sex models in comparison with divergence for the age-structuredsingle-sex models (Section~\ref{sex:doublingex}). Finally, inSection~\ref{sec:exdecomposer} we quantify the vital rate components to thesex-gap in the thanatological $r$ using a flexible method of demographicdecomposition. This closes our treatment of single-sex renewal for populationsstructured by remaining years of life, leaving much terrain unexplored.Having developed a working single-sex model for the remaining-years perspective,Part 3 of this dissertation translates and explores a set of the two-sexmethods from Chapter~\ref{sec:modelingapproaches} to thanatologicallystructured populations. This is carried out in much the same sequence as wasdone for age-strctured two-sex models, first describing the fertilitybalancing method and deriving the renewal model, second explaining step-by-stepinstructions for estimating $r$ (and SRB it turns out) from given data, thirdworking out a selection of some other stable parameters, andfinally a demonstration of (some aspect of) the method on the basis of the USand Spanish data used throughout this dissertation.We begin the two-sex methods translation with thedominance-weighted method of \citet{goodman1967age} inChapter~\ref{sec:ex2sexdomweights}. As this is the first two-sex renewal modelexplored for remaining-years populations, we take some extra care to compareresults both with the remaining-years single-sex models and with age-structured models. Notably, wedefine the two-sex dominance-weighted remaining-years structuredprojection matrix (Section~\ref{sec:ex2sxprojmat}), and use this to exploreseveral aspects of the stable population structure(Section~\ref{sec:ex2sexdomweightsstabstruct}). Second, we translate themethod based on a generalized mean of male and femaleremaining-years specific exposures, also for the most part on the basis of theharmonic mean (Chapter~\ref{sec:ex2sexschoen}), although results do not varymuch if other common (and reasonable) means are chosen instead. The values ofthe intrinsic growth rate, $r$, that one derives with this method are very closeto those given by the dominance-weighted method when equal weight is given tothe male and female rates. For this reason we focus the empiricaldemonstration of the method on other stable consequences that do not apply inthe latter case, most notably on the stable versus the initial $e$SFR.Third, in Chapter~\ref{sec:CRchap} we sketch out a novel two-sex adjustmentmethod for this population structure based on marginal male and female fertilityrates and a fixed ratio of observed birth counts to association-free expectedbirth counts in the joint male-female distribution. The model is seen to giveacceptable results, but we conclude that it displays no particular advantageover the generalized mean or IPF methods, and so we discard it in order to moveon to the IPF method.Finally, in Section~\ref{sec:ipfex} we translate the iterative matrix methodoriginally presented in Section~\ref{sec:IPF} to the two-sex remaining-yearsstructured population. This method indeed yields estimates of $r$ that are distinct fromany of the previous methods (for at least one of our two populations). We thenexamine the initial versus stable fertility rates, and compare results withthose obtained from the generalized mean method.That the methods and observations presented here are so out of the ordinaryinvites one to reflect rather than to conclude, although inChapter~\ref{sec:reflections} we attempt both. Namely, we consider whetherthere is anything to be gained by conceiving of reproduction in terms of remaining years of life, why it is that theremaining-years structured family of models yields a different estimate ofpopulation growth than does the age-structured family of models, and how ourexperience with the two-sex problem might yield insight on this lack ofcongruence. Further, we summarize any empirical findings of note, though theseare not central to our objectives. Finally we summarize theoreticalcontributions from this work and suggest a broad and ambitious research agendato be grounded in it.\section{Data}\input{sections/other/Data.tex}