 \FloatBarrier
\label{sec:modellingapproaches}
\begin{singlespace}
\begin{quote}
The purpose of models is not to fit data but to sharpen questions-- Samuel
Karlin\footnote{Quote from the \nth{11} annual R. A. Fisher Memorial Lecture
given at the Royal Society of London on April \nth{20}, 1983.}
\end{quote}
\end{singlespace}

The an assessment of the magnitude and vital rate components to the two-sex
problem in age-structured populations was presented in the previous
chapter for the cases of the US and Spanish populations. That a fundamental
discrepancy exists between the male and female single-sex models was already
empirically demonstrated by \citet{kuczynski1932fertility} in the early 1930s, and formally introduced to
the discipline of demography in the late 1940s by
\citet{karmel1947relations, karmel1948measurement}, since which time interest
has continued in waves. It is the case that the discrepancy thus far has no real 
solution in the sense of a necessarily true
solution. Instead what is meant by a two-sex \textit{solution} is a method to
balance male and female vital rates so as to produce the same estimate of births (or marriages), 
or else the same structured stable population. This is
perhaps part of the reason why interest has continued-- there are many ways in which this goal
might be achieved. All methods proposed have incurred some degree of trade-off
between convenience, simplicity, realistic design and results that are
consistent with expectations.
 
The point of departure for the goal in formal demography of balancing male and
female rates is the following: For the single-sex case-- the classic Lotka
model captured in Equation~\eqref{eq:lotkaeq}-- we have a coherent model that
works for each sex separately but produces undesirable results when modelled in parallel for both
sexes. What modifications must we introduce to the model, such that a single
estimate of the intrinsic growth rate, $r$, is produced while maintaining a
reasonable sex ratio (both total sex ratio and the sex ratio at birth) and
maintaining constant male and female vital rates? 

An alternative formulation
could be summarized in terms of producing a single prediction of births in
future projected years. Namely, what changes should we admit to the cohort
component projection method such that the model accepts both male and
female inputs but produces consistent output in the form of single estimates of
male and female births.

Classifying two-sex models into families of models that share similar qualities
is non-trivial. We will present an imperfectly arranged subset of models that
have thus far appeared in the literature, focusing primarily on those models
related to those that we will modify later in this dissertation for the case of
population structured by remaining years of life. These include models amenable
to using single-age discrete aggregate data, that do not require a global
optimization of a large number of parameters, and that do not include marital
or other states. In other words, the models presented here exclude most recent
advances in the two-sex modelling per se, which have been in the areas of agent 
based models and differential equations. That is to say, we do not delve into the 
interior workings of mating or marriage markets or individual interactions, or
into the epidemiological or econometric interests in the two-sex problem in general. Work
is kept in the domain of classical aggregate demography, and models easily
summarized in terms of unified renewal equations. In this sense the work of
this chapter is not \textit{cutting-edge}, but rather one of standardizing, at
time translating models from their original context to fit ours. In all cases,
we attempt to give step-by-step instructions for how to calculate the
two-sex intrinsic growth rate, $r$, to ensure reproducibility of results.

We begin by classifying into a single group those two-sex models whose
constituent parts have essentially been the male and female single-sex models;
i.e. those models that have adhered to the concept of single-sex fertility. In this group
we could first place the exceedingly simple solution that consists in taking some 
mean of the male and female single-sex intrinsic growth rates to produce a 
both-sex $r$, without digging into the workings of the model
itself\citep[e.g.][]{kuczynski1932fertility}. We will discuss two models that
fall into this class, one parsimonious and effective, another intricate. Both
models will be seen to yield results in line with expectations, but to suffer
particular drawbacks. The first model which we will briefly discuss is that
which appeared in \citet{pollard1948measurement}. This will be followed by
a longer explanation of the less intuitive model in \citet{mitra1978derivation}.
Neither of these two models will later be translated to the remaining years
perspective.

Next we briefly explain a linear model that makes use of a constant
\textit{dominance} parameter--a weighted mean-- to regulate the
influence of male and female vital rates on intrinsic growth rates. In this
model, both males and females produce offspring of both sexes, thereby
removing any additional complications in sex ratios implied by the use of
single-sex fertility rates. This model is later translated to the
remaining-years perspective later in this dissertation.

Third, we present two more models, \citet{gupta1978alternative} and
\citet{schoen1981harmonic} (and mention several others) that make use of
fertility rates cross-classified by age of mother and age of father. Fourth, 
we present an iterative model that also relies on age
cross-classified fertility information, but that has some more desirable
properties, namely the iterative proportional fitting technique first
introduced to the two-sex problem by \citet{mc1975models}. Some results are
 compared and assessed in light of the axioms presented previously. Finally, a
 brief survey of methods not covered at length will close the chapter.
 
\FloatBarrier