 \FloatBarrier \label{sec:axioms}It has been pointed out that the ideal functional form of a two-sex solutioncannot be empirically determined (Das Gupta, Schoen, Keyfitz, others). This isbecause fertility is always undergoing secular changes, to the effect thatone cannot simply calibrate an ideal mean function (if a mean function were thecorrect choice) net of outright both-sex fertility change. This we observe aboveall with the Spanish data used in this dissertation: from 1975 until the mid1990s fertility levels dropped so rapidly that in most cases the year $t+1$birth count fell below that which would have been predicted by either of theyear $t$ male or female rates-- despite there having been a wide the gap betweenmale and female total fertility rates in those years. Even in lessextreme situations, where the year $t+1$ birth count is intermediate to that which wouldhave been predicted by the male and female year $t$ rates, one is unable toseparate the effects of relative changes in male versus female exposure fromsimple changes in rates. That is to say, if there is some push and pull betweenmale and female rates, this cannot be measured if rates on the whole are eitherrising or falling-- just as it is difficult to measure the net rising andfalling of rates when there is feedback and separation between male and femalerates. Even if one had a very large amount of data conformable to this problemand an appropriate statistical technique so as to mete out these differences and estimate a function that could separate and capture the effects of ourimagined push and pull between male and femalerates\footnote{\citet{alho2000competing} come close to this ideal.}, it would be easy to suppose that this ideal functionmight itself change according to certain conditions or certain periods.This empirical obstacle has led demographers to devise a set of axioms,necessary or desired characteristics, that the ideal two-sex fertility(or marriage) function should abide by in order that it conform with ourexpectations. Here we will enumerate all such axioms located in the literature before briefly discussing them in turn. Here, $M()$ is any function thatdetermines the both-sex rate using male, $P^m$, and female, $P^f$, exposures asinputs. These exposures may be classified by some other variable, such as age, butsubscripts are ignored here unless pertinent.\begin{description}  \item[Availability:] $M(P^m, P^f) = 0$ if $P^m = 0$ or $P^f = 0$. Members of  both sexes must be present in order for there to be a non-zero positive rate.  \item[Homogeneity:] $kM(P^m, P^f) = M(kP^m, kP^f)$. Equal  changes in the supply of males and females must lead to an equal change in the  number of births (marriages).  \item[Monotonicity:] for $k > 1$, $M(kP^m, P^f) \ge M(P^m, P^f)$ (and vice  versa). If the supply of one sex increases while the other sex is held constant, the number of  births (marriages) cannot decrease.  \item[Symmetry:] for $P^m = P^f$, $M(kP^m, P^f) = M(P^m, kP^f)$.   \item[Competition:] if exposure in age $x$ for males is increased by some  factor, but all other male and female ages are held constant, monotonicity  applies to age $x$ of males, but rates for male ages $<x$ or $>x$ may only  decrease or say the same.   \item[Subsitution:] The size of competition effects varies directly with  age-proximity to $x$ among males. For instance, males of age 24 are closer  substitutes for males of age 25 than are males of age 20.  \item[Bracketing:] $M(P^m, P^f) > min(F^m, F^f)$ and $M(P^m, P^f) < max(F^m,  F^f)$. The both-sex rate must be intermediate to the single-sex rates.  \item[Proportionality in the extreme:] in situations of very extreme sex-ratio  imbalance, changes in the amount of the minority sex should be reflected proportionately  in the two-sex rate.\end{description} These axioms will now be briefly reflected upon in turn.\paragraph{Availability:} This is the most elemental axiom, as it essentiallystates a truism: If one sex is absent, there can be no reproduction in a speciesthat reproduces sexually. For the sake of philosophical completeness, we statethe following: 1) Assisted reproduction requires both sexes, so this is noretort; 2) At present, technology that would negate this axiom, humanparthenogenesis, is not fully developed, although ther have been recentadvances\citep{revazova2007patient}. If and when technology would permit asexualhuman reproduction, there will be legal hurdles, costs, and apoption lag. That is to say, potential anecdotes that would negate this axiomwill in any case not affect fertility rates in a significant way within the timehorizons that demographers currently project. This is not a tongue-in-cheekobservation, as technology in general is known to affect fertility in myriadways. For instance, in vitro fertilization and other forms of assistedreproduction have had noticeable effects in the fertility and sex ratio at birth from particularage groups.\paragraph{Homogeneity:} This author finds the axiom of homogeneity to be on thewhole harmless, but not necessarily true. Homogeneity essentially states thatthere are no scaling effects. It is easy to imagine that population size willconstrain or determine much of what happens within populations. This isespecially so when we think in terms of social organization, contactopportunities and the countless other structural factors that may affect thepractice of mating and by extension fertility. Population size has beengiven more attention in non-human ecology \citep{donalson1999population} than inhuman demography, where considerations of population size have been primarily framed in terms ofcarrying-capacities \citep[see e.g.][]{cohen1995human,hopfenberg2003human}. Thisauthor is only aware of scaling in demographic process when studied as complex systems viaagent based modelling (ABM) \citep[e.g.][]{bruch2010scaling}. While ABMs havebeen used to studying fertility and marriage \citep{billari2002wedding},indirect scaling effects in such models have not been explicitly studied, nothave scaling effects been in introduced explicitly.\paragraph{Monotonicity:} This may seem intuitive, but if viewed from asociological standpoint, it seems plausible that increased competition could actually lead to adecrease in total births (marriages) via different mechansisms that we willbriefly hash out. Imagine a more complex model wherein individuals mustapportion time / effort / resources between mate search costs and competition. In the case of an increase in males while holding femalesconstant, increased competition between males in mate selection could scale non-linearly to-- and offset-- the standard predicted increase in matings that would result from increased male pressure on the market. In a different scenario, females faced with abundant potential mates may actually decrease their  search efforts and postpone the mate search until a later time, thereby acting to supress rates. Were this later effect present in the model, the effect of increases in one sex would be ambiguous, as it would depend on the relative forces of male pressure and female deprioritization. In yet another model scenario, a proportion of males faced with increased competition may  indeed cease to compete, and remove themselves from the market, thereby  decreasing pressure from the side of abundant males. Other similar effects may be dreamed up wherein the results of an increase in one sex only could be coplex and counterintuitive.   None of these complex model scenarios are particularly amenable to inclusion in a practical analytic model of mating / marriage / fertility markets. However, in indicating such potential countervailing forces-- all reasonable in the mind of this author-- one may at least question the necessity of holding monotonicity as \textit{axiomatic} in the sense of a functional necessity- a criterion by which the adequacy of a model may be judged. \paragraph{Symmetry:} It appears that symmetry, treated as an axiom, is alsoinappropriate. Males and females differ not only with respect to vital rates,but with respect to mate preferences and behavior \citep{buss1989sex}. There isalso evidence for variation in these differences by group size\citep{fisman2006gender}, which plays into the previous axiom of monotonicity.Clearly, if males and females have different preferences and also reactdifferently to differences in group size, we should expect different outcomesfrom hypothetical sex-complementary compositional changes in the mating market.For this reason we may also conclude that symmetry, though likely to be acharacteristic of the functional form assigned to the male-femaledependant fertility (marriage) function, ought not be given the status of anaxiomatic requirement for a good and proper model. That the functional formsoften used for marriage and fertlity often were symmetric with respect to thesexes need not be drawback, but we ought not grant this characteristic post hocstatus as an axiom.\paragraph{Competition:} It seems reasonable that, holding mate supply constant,increases in matings in age $x$ either decrease or have no effect on agesclose to $x$ of the same sex. Some two-sex models have accounted for this axiom\citep{henry1972nuptiality, mc1975models, choo2006estimating}, sometimes viaexplicit preference functions \citep{parlett1972can, pollard1993interaction} butmany have not. These models are considerably more complex to implement than the alternatives. It is unclear to this author whether this axiom should be treatedas a requirement or a desirable property. \paragraph{Substitution:} In the case of inter-age (or inter-group) competitionfor mates, it is intuitive that, since age can be thought of ascontinuous, competition effects should vary inversely in magnitude as a functionof distance to the age that experienced a sudden change in effectivepopulation. In the case that explicit preference functions are used, this axiomis already dealt with, and \citet{choo2006estimating} also has thischaracteristic.\paragraph{Bracketing:} The interpretation of this axiom depends on context.In the first instance, it states that the two-sex instrinsic growth rate, $r$,must fall between the male and female single-sex intrinsic growth rates, $r^m$, $r^f$, respectively. Many authors have treated this axiom as a requirement\citep{pollard1948measurement,coale1972growth, gupta1976interactive, mitra1978derivation}, others haveargued otherwise \citep{gupta1973,schoen1981harmonic}, and indeed it has evenbeen proven an unreasonable condition \citep{yellin1977comparison}. This authoragrees that the single-sex growth parameters will not serve as two-sex boundsbecause they are calculated in unreasonable isolation, namely, each constrainedby its own sex-specific fertility rates and without interaction between ages ofeach sex. That is to say, in isolation the single-sex models may behavestrangely and not bound the true trajectory of the total population.A second domain of bracketing could be in terms of the total births predicted bymales and females for year $t+1$ using the ASFR and sex-specific exposures. Inthis case, the main difference is that the offspring of each sex is of bothsexes. In this case, bracketing appears a less troublesome condition, as weessentially remove fertility sex-ratio constraints from the boundarypredictions. Absent secular change in birth rates or the age-pattern of fertility, we would expect one sex to ovestimate and the other to underetimate the birth count to be observed infuture years. \paragraph{Proportionality in the extreme:} In other words, at some point alongthe continuum of potential sex ratios, the minority sex should experience\textit{saturation}, in the sense that further increases in the majority willnot result in increased matings. In this same scenario, one may imagine that,while still within the same extreme order of sex ratio magnitude, a unitincrease in effective population of the minority sex will lead to a unitincrease in predicted births (marriages). It is doubtful that thissituation would ever arise in a real projective scenario 