This author has provided no proof that the stable rate of growth
attained in the various two-sex age-structured and remaining-years structured
models are unique, necessary or ergodic. There is a possibility that under some
real conditions the stabilizing trajectory arrives in a limit cycle, which this
author considers a particular variety of stability. \citet{wijewickrema1980weak}
and \citet{chung1990phd, chung1994cycles} explore the possibility of such cycles
and bifurcations in two-sex models, but this and many other dynamic
properties remain to be explored for remaining-years structured two-sex
populations. We have omitted any sensitivity analysis, although this would
enhance our ability to compare age-classified and remaining-years
classified models. We have in some cases measured the total amount of
oscillation in population structure between the initial and stable states, but
we have not examined its path. We have also not checked whether the end state
is truly ergodic or somehow dependent on initial conditions. These and most
other transient properties of the models presented in this dissertation have
been ignored, and are particularly ripe for exploration for the new family of
remaining-years structured models that we propose.

Assuming that the patterns to fertility by remaining years of life
are indeed meaningful, and in any case observing that they are regular (for
some, the only requisite), we are now free to model population on the basis of
them, just as demographers have always done on the basis of age. This we began
for the single-sex case in Chapter~\ref{sec:exstructuredrenewal}, and there we
learned in the first place that the instrinsic growth rate that belongs to this
\textit{family} of model is less erratic than the age-analogue. In other
contexts, demographers have lent value to less-erratic renditions over otherwise
familiar demographic time-series. Such has been the case with tempo-adjusted
fertility rates \citep{bongaarts1998quantum}\footnote{Although authors
\citep[e.g.][]{kim2000quantum} have found this species of adjusted TFR to be
erratic as well, the basic desire for a relatively stable indicator remains.}.
Here we have produced such a series for instrinsic growth rates (and $R_0$ and $e$SFR), 
more stable than the chrononological-age analogue, and the reader must be the 
judge of whether this adds value. We expect that many demographers would prefer
to tinker before passing judgement. For this reason, we make available the 
basic transformation of Equation~\eqref{eq:dxredist} in both a spreadsheet and
an \texttt{R} function.

As for why results of chronological age should differ from those of
thanatological age, we noted above that the pertinent rates are calculated on
the basis of different underlying exposures. There is room to experiment with
 finer segmentation of exposures, but we are uncertain (but not doubtful)
 whether anything is to be gained by a more complex model. Clearly one can
coherently conceive of population renewal under the remaining-years
perspective, and we have built models that accord with this vision. A parallel
may be drawn with male and female single-sex models under either variety of
age; both models are of equal soundness, yet return results that are at odds. 
It is tautologically the case that in the year of initial conditions, the male and
female models produce no discrepancy. We may say the same of models under
thanatological and chronological age; in the intitial year these two models are
congruent, but thereafter they diverge.

With the two-sex problem, it is easy to imagine that the two sexes modelled
separately are bound to diverge, and to note that this may not be so in
observerved populations, as governed by the sex ratio at birth. Any two-sex
model will provide that the two sexes project through time in unison. Model
divergence under chrononological versus thanatological age, on the other hand,
will result even when both sex sexes are modelled together under the same principles under each
of the two age structures. Two sex models for either age definition are a
summary of the growth of the entire population, yet results depend on whether
one counts up from birth or down toward death. We are happy to have demonstrated
this discrepancy, but this finding is rather unglamorous in comparison with a
hypothetical model that would contain information from both age definitions and
both sexes, or else a philosophical argument for why one definition of age
produces a superior model of population growth. Clearly a population may only
have one total growth rate or net reproduction ratio. Let us call this cunundrum 
for now the \textit{two-age} problem.

While the two-sex problem has not been solved in an necessary and true way, some
satisfactory solutions have arisen. These solutions have in common that they
deal somehow with mixing, with the interaction between sexes, and axioms have
been developed to help guide the way in determining ideal model properties. No
such axioms exist at this time for the two-age problem-- This author does not
even know how to properly frame it. No model has been proposed
that would unify the results of these two definitions of age. Recall that some 
of the initial responses to the pointing-out of the two-sex problem
were to produce ad hoc justifications for female dominance, and some of these
have had staying-power. These issues are worthy of more contemplation than that.
After \citet{karmel1947relations} formal demographers came to realize the
importance of modelling the two sexes together, and a great body 
of work has been produced to this end. 