
Population projections rarely incorporate males into fertility assumptions.
 In Chapter~\ref{ch:Measuring} we showed that this \textit{female dominance} is
 not an innocuous assumption for the case of age-structured models, and in
Chapter~\ref{sec:exstructuredrenewal} we further showed it is also
not an innocuous assumption for the case of the remaining-years structured
populations proposed here. In Chapter~\ref{sec:modelingapproaches} for
chronological age and the chapters of Part 3 for thanatological age, we set out to provide a suite of
standard working examples of solutions of two-sex adjustment methods so that 
these can be compared and implemented by the demographer if desired. This 
is a practical concern that has implications for how
contemporary demography is practiced. In this section we provide some
guidance by recommending from among the methods presented.

The two-sex problem comes to the fore when projecting
population into the future. The most common practice to avoid disagreement between the sexes is to
assume female dominance, which characterizes the female single-sex Lotka
renewal model, and is a degenerate case of the
\citet{goodman1967age} two-sex model from Section~\ref{sec:googmanage} with the 
dominance parameter set to accept 100\% of model information from female
fertility rates. There are very few population projections produced for 
consumption beyond two-sex studies themselves that do not make this assumption. 
Assuming female dominance in one way or another is so widespread that newcomers
to demography often perceive it as a given, or in the worst case adopt the
practice dogmatically. For others, the implementation is too 
complex or the requisite transition rates are unavailable. These latter
two obstacles can be overcome by omitting nuptial states from the model
altogether, as we have done in this dissertation. 

Formal demographers have long taken sex-divergence seriously, at
least for purposes of consistent model design. These mathematical models
have been the driving force behind this dissertation effort. We have at
times simplified two-sex models from their original form, and one offshoot of
this choice is that our palette of models might be of potential use to applied
demographers. The \texttt{R} code used to produce our results should be
recyclable, or else can easily be quarried to such ends. It therefore behooves
us to recommend from among the methods explored here. In order to account for
sex divergence in projections or self-contained models, of those methods treated in this dissertation, 
we recommend from the following three adjustment strategies.

\begin{enumerate}
\item The weighted-dominance method of \citet{goodman1967age} from
Section~\ref{sec:googmanage} and Chapter~\ref{sec:ex2sexdomweights} is a
reasonable choice, probably with a dominance parameter of $0.5$, 
such that equal information is taken from male and female
weights. This method wins on parsimony and ease of implementation, and it has
the simplest data inputs. It produces reasonable results for population
structures typically observed (i.e., without zeros in reproductive ages). It is
also containable in a static Leslie matrix, although
we  explain this construction only for populations structured by
thanatological age (see Section~\ref{sec:ex1sexleslie}). The method has a
drawback in that the model itself is less appealing, as it does not allow 
for proper interaction between sexes, or ages. However, for the purposes of
projecting populations structured by age and sex within a 40-year
horizon\footnote{This is an arbitrary threshold. The idea is that in the near
term results between models do not differ by much.}, the simplest model is well
worth consideration.

\item The mean method from Sections~\ref{sec:ageharmonic} and
\ref{sec:ex2sexschoen} for age and remaining-years structured models
respectively, is also a reasonable choice. In this case, we recommend
implementation with the mean set to either harmonic, logarithmic, geometric or
an unnamed general mean in that approximate range (with the property of 
falling to zero if one sex is absent). There is no point in
quibbling over which of these is best, as observed human populations do
not exhibit the extreme sex ratios required to produce meaningfully
different results. If one were to further segment the population into
categories on which partner preference occurs, then more extreme sex ratios
would be possible -- but then the whole model design also becomes more complex.
This method is appealing because the male and female marginal fertility rates 
for a given year are determined dynamically by changes in each 
age-combination (remaining-years combination) of males and females, and the
range of means listed here allows for some degree of \textit{bottlenecking} due
to the minority sex in a particular combination. The model is also of
parsimonious design, easy to understand, and straightforward to implement.
However, this model does not allow for competition or substitution between ages
(remaining years), which could be a drawback if the demographer wishes to
infer future age-specific fertility rates through abrupt changes in cohort size.

\item The iterative proportional fitting method (IPF) is the most flexible of
these, because it does incorporate competition and substitution between ages
(remaining years). That the
method is iterative presents no real drawback, as any of the above methods is best 
implemented in a scripted language, and the method is
not perceptibly slower to calculate than the alternatives. The properties of
IPF are easy to demonstrate but difficult to prove, and so it has received less attention from
mathematicians and continuous-equation modelers. However, for the demographer
designing a projection apparatus, IPF is nonetheless a convenient choice. This method is in need of
comparison with the recent contribution from \citet{choo2006estimating}, which
has similar properties, but which we do not discuss here.
\end{enumerate}

Each of these three methods has advantages and disadvantages, and it is up to
the demographer to evaluate the optimal choice for a particular projective or
modeling scenario. All three work equally well in
age-structured and remaining-years structured projections. Further, these
adjustment techniques work just the same for projections that incorporate
assumptions about future developments in fertility. In this case, the
demographer makes assumptions about the male and female paths of fertility-rate development and adjusts in each
iteration to force agreement in results. For the mean method, the incorporation of sophisticated assumptions entails more care, as these must be
distributed over a matrix. If this proposition is inconvenient, one might prefer
either IPF or dominance weighting, which rely only on marginal fertility
distributions for adjustments beyond the initial year.

Further, we have demonstrated that mean-based and IPF methods
entail changes to the marginal male and female rates after adjustment, and one
might wonder whether these adjusted rates have any predictive power per se --
for instance, adequately adjusting for foreseen changes in population structure
or whether they are a modeling artifact to be disregarded in favor of the total (unstructured) birth count. We have demonstrated that this feature exists,
and we have shown instances where the two methods make predictions that are at
odds. We do not follow up this observation with an empirical comparison
to determine which hits the mark closer in terms of
fertility distribution prediction. This therefore remains an intriguing question
(which has been ignored thus far in the literature) that could tip the
balance in favor of one of these methods. A priori, we expect IPF 
to display more appropriate sensitivity amidst abrupt
changes in cohort size, but we do not know whether the magnitude and
distribution of adjustment is appropriate. 

Of course, the demographer can also consider two-sex models not treated
in this dissertation, perhaps incorporating nuptial states into the model. 
In this case, the two-sex method is transferred to nuptiality --
match-making, pairing, marriage -- as the event being predicted, but the
adjustment procedures are one and the same. Such a projection
entails more sophisticated construction, more data inputs, and the incorporation
of more hypotheses, namely hypotheses (specifically regarding changes in
marriage rates, as well as changes in marital and extramarital fertility). For
populations with high proportions of extramarital fertility, extra data are required to approximate the formation of
non-marital mated pairs -- for example, transitions into and out of
cohabitation, as well as fertility rates that apply to this subpopulation (and mortality
rates if supposed different). That is to say, adding further state
considerations to fertility assumptions greatly increases model complexity
and data requirements, and so this is not likely to be appealing to projection
designers unless predictions of the states themselves are necessary or
inherently interesting.

To the extent that fertility rates and the sex ratio at
birth vary along the path to stability, one might wonder whether any of the
\textit{interactive} two sex models are at odds with the notion of
rate invariance in stable populations. In any of the
interactive models, the element held fixed prior to stability is not rates, but
some standard. For IPF, the element held constant in our 
description is the original cross-classified birth matrix and corresponding 
male and female marginal rates. For the mean-based method, one holds constant 
the standard rate matrix, as well as the mean function itself, but the marginal 
rates produced by these standards have been shown to change over time
under these modeling assumptions. Once in the state of stability, of
course, both population structure and marginal male and female fertility rates are invariant, which
implies that the two-sex problem itself vanishes. In this case, for both the
mean and IPF methods, the stable adjusted marginal fertility rates become
invariant, and the male and female rates yield the same results -- making the
population tautologically \textit{dominance-indifferent} once in stability.

These recommendations are made not for the sake of modeling, but for the sake of
applied demography: projections. The first two of these recommendations
violate at least one axiom, and the last (though easy to program), is a sort of
black box, and it is difficult to describe in a parsimonious way.
